---
title: "Covid91 vaccine study Final2021"
author: "Tania Boror"
date: "5/4/2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

#  Methods 

The first step is isolating the subpopulations of Males, Females, LGBTQ, and DrugUsers.


# MALES

```{r}
Maledata <- filter(FinalData,sex== "guy")
Maledata$sex <- as.factor(Maledata$sex)
Maledata$infected=factor(Maledata$infected)
Maledata$infected<-relevel(Maledata$infected, "OK")
```

## Bar Chart

```{r}
Males <- Maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately 340 Males who were infected with Covid-19, 57% were given the drug, and 43% were given the placebo.

##  Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=Maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

The table shows that about 340 males were infected with Covid-19 out of approximately 10,400 males in the study. The table shows that 3.72 % of males who receive the drug got Covid 19, and 2.55% of the males who got the placebo contracted Covid-19. The table also shows, that males that got the vaccine were about 50% more likely to contract Covid-19, this suggest that the vaccine is not effective for males.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Squared table shows that the P-Value of 0.0006181 is so small that it is highly unlikely that the results are by chance.  
$H_0$ the null hypothesis given taking the vaccine there is no difference of whether males can contract Covid-19 or not.  
$H_A$ the alternative hypothesis is that the vaccine in fact does make a difference in whether or not a male contracts Covid-19. Therefore we Reject the Null Hypothesis.

The odds ratio from the Fisher Exact test is 0.68:1 and the confidence interval is between 0.54 - 0.85, which is less than 1. This means that the effectiveness is .68/(1+.68)*100, which is 40.5% effective. 

# FEMALES

```{r}
Femaledata<- filter(FinalData,sex=="gal")
Femaledata$sex <- as.factor(Femaledata$sex)
Femaledata$infected=factor(Femaledata$infected)
Femaledata$infected<-relevel(Femaledata$infected, "OK")
```

## Bar Chart

```{r}
Females <- Femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately 607 Females who were infected with Covid-19, 32% were given the drug, and 66% were given the placebo. The chart shows that the drug is effective against covid-19. 


##  Numerical Summary

```{r}
table2 <- xtabs(~treatment + infected, data=Femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```

The table shows that about 607 Females were infected with Covid-19 out of approximately 9,200 Females in the study. The table also shows that 4.44 % of Females who receive the drug got Covid-19, and 8.67% of the Females who got the placebo contracted Covid-19. Lastly the table shows, that Females that got the vaccine were about 67% less likely to contract Covid-19, this shows that the vaccine is effective for Females.

## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The Chi Squared table shows that the P-Value of 0.0000000000000002 is so small that it is highly unlikely that the results are by chance.  

$H_0$ the null hypothesis is that the vaccine makes no difference in the female population when it comes to contracting Covid-19. 

$H_A$ the alternative hypothesis is that the vaccine in fact does make a difference in whether or not a Female contracts Covid-19. 

Therefore we Reject the Null Hypothesis.

The Fisher Exact Test odds ratio 2.05:1, when taking the odds radio and computing it into a percent of effectiveness by using the formula O/(1+O) * 100 you get 67% effectiveness.

Ultimately, the vaccine is 67% effective in the Female population, which is absolutely effective.

# DRUG USERS

```{r}
Druggiedata <- filter(FinalData,DrugUser== "yes")
Druggiedata$DrugUser <- as.factor(Druggiedata$DrugUser)
Druggiedata$infected=factor(Druggiedata$infected)
Druggiedata$infected<-relevel(Druggiedata$infected, "OK")
```

## Bar Chart

```{r}
Druggies <- Druggiedata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately 117 DrugUsers who were infected with Covid-19, 32% were given the drug, and 66% were given the placebo. The drug is effective for those who are drug users.

##  Numerical Summary

```{r}
table3 <- xtabs(~treatment + infected, data=Druggiedata)
rowPerc(table3)
colPerc(table3)
print(table3)
```

The table shows that about 117 DrugUsers were infected with Covid-19 out of approximately 2,280 DrugUsers in the study. The table also shows that 33.3% of DrugUsers who receive the drug got Covid-19, and 66.6% of the DrugUsers who received the placebo contracted Covid-19. Lastly the table shows, that DrugUsers that got the vaccine were about 60% less likely to contract Covid- 19, this shows that the vaccine is effective for DrugUsers.

## Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The Chi Squared table shows that the P-Value of 0.002273 is so small that it is highly unlikely that the results are by chance.  

$H_0$ the null hypothesis is that the vaccine makes no difference in the DrugUser population when it comes to contracting Covid-19. 

$H_A$ the alternative hypothesis is that the vaccine in fact does make a difference in whether or not a DrugUsers contract Covid-19. 

Therefore, we Reject the Null Hypothesis.

The Fisher Exact Test odds ratio 1.83:1, when taking the odds radio and computing it into a percent of effectiveness by using the formula O/(1+O) * 100 you get 60% effectiveness.

Ultimately, the vaccine is 60% effective in the DrugUsers population, which is effective.

# LGBTQ

```{r}
LGBTQdata <- filter(FinalData,LGBTQ== "gay")
LGBTQdata$LGBTQ <- as.factor(LGBTQdata$LGBTQ)
LGBTQdata$infected=factor(LGBTQdata$infected)
LGBTQdata$infected<-relevel(LGBTQdata$infected, "OK")
```

## Bar Chart

```{r}
LGBTQ <- LGBTQdata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately 116 LGBTQ who were infected with Covid-19, 70% were given the drug, and 29% were given the placebo. 

##  Numerical Summary

```{r}
table4 <- xtabs(~treatment + infected, data=LGBTQdata)
rowPerc(table4)
colPerc(table4)
print(table4)
```

The table shows that about 116 LGBTQ were infected with Covid-19 out of approximately 2,040 LGBTQ in the study. The table also shows that 70% of LGBTQ who receive the drug got Covid-19, and 29% of the LGBTQ who received the placebo contracted Covid 19. Lastly the table shows, that LGBTQ that got the vaccine were about 29% more likely to contract Covid-19, this shows that the vaccine is not effective for the LGBTQ population.

## Inferential Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The Chi Squared table shows that the P-Value of 0.0000057 is so small that it is highly unlikely that the results are by chance.  

$H_0$ the null hypothesis is that the vaccine makes no difference in the LGBTQ population when it comes to contracting Covid-19. 

$H_A$ the alternative hypothesis is that the vaccine in fact does make a difference in whether or not a LGBTQ contract Covid-19. 

Therefore, we Reject the Null Hypothesis.

The Fisher Exact Test odds ratio 0.40:1, when taking the odds radio and computing it into a percent of effectiveness by using the formula O/(1+O) * 100 you get 60% effectiveness.

Ultimately, the vaccine is 29% effective in the LGBTQ population, which is absolutely ineffective.

# Discussion & Conclusion

When looking at the 4 populations, we found that the two populations that had the highest rate of effectiveness were the Females at 67% and the DrugUsers at 60%. Both the Male population and the LGBTQ populations showed that the vaccine was not beneficial to them and were ineffective in protecting from contracting Covid-19, the Males had an 40% effectiveness, while the LGBTQ had a 29% effectiveness, which is not effective.
